
Week 4 Project
=============================================================

***Data Set Description***

Data set file name: World_Bank_Indicators.csv
Source: http://data.worldbank.org/indicator

This data set contains 3 indicators from the World Bank on Mortality Rates and Economic factors for every country in the world between 2000 and 2010. 

A complimentary dataset provide categorization of every country by region and subregion.

***Data Structure***

**Dataset: World_Bank_Indicators.csv***

One instance per line with comma delimited fields.

No. of Observations: 2354

No. of Variables: 5

**Variable name::          Description::	                           Data type**

Country.Name::	        Country Name::	                           Char

Date:: Effective date of indicator:: Date

H.Mortality.U5::	        Mortality under 5 yo per 1000 ppl::	      Number

H.Exp.PerCapita ::       Health expenditures per capita::	         Number

F.GDP.PerCapita::	GDP per capita in USD::	Number

**Dataset: regions.csv**

One instance per line with comma delimited fields.

No. of Observations: 234

No. of Variables: 3

**Variable name::   Description	Data type::	Valid values**

Region::	Region name::	Char	

Subregion::	Subregion name::	Char	

Country::	Country name:: 	Char	


***Data Analysis***


**Data Cleanup**

1. First data load loaded only 512 record. Upon research it was found that a comma in the country name Cote d'Ivoire must be escaped before complete data set can be loaded

2. When all rows were loaded into a data frame it was found that most of the numeric fields were converted to factors. This happened because numbers were formatted with a comma separator as follows: 8,996,544. After removing commas all numbers loaded correctly as integers.


**Load required packages**

```{r}
library(ggplot2)
```

**Load data**

```{r}
setwd("C:/Igor/CUNY/607_Data_Acq_Mgmt_R/Week_05/project")
ind <- read.table(file = "world_bank_1.csv", header = TRUE, sep = ",")
reg <- read.table(file = "regions.csv", header = TRUE, sep = ",")
str(ind)
str(reg)
```

Join two datasets in order to be able to analyze data by region and subregion

```{r}
wbi<-merge(x=ind,y=reg,by.x=c("Country.Name"),by.y=c("Country"))
```

Summary of all columns in the data set:

```{r}
summary(wbi)
```

Summary shows that all fields are within acceptable ranges with no apparent outliers

We will analyze latest data for 2010. Let's see if our data proove a common believe that countries with higher Health Expenditures and higher GDP Per Capita have lower Children Mortality Rate.

```{r}
wbi$Date2 <- as.character(wbi$Date)
wbi2010<-wbi[wbi$Date2==("7/1/2010"),]
options(scipen=5)
qplot(H.Mortality.U5, H.Exp.PerCapita, data = wbi2010,xlab="Health Expenditures Per Capita", ylab="Children Under 5 Mortality Rate", main="Health Expenditures vs Mortality Rate")
qplot(H.Mortality.U5, F.GDP.PerCapita, data = wbi2010,xlab="GDP Per Capita", ylab="Children Under 5 Mortality Rate", main="GDP vs Mortality Rate")
```

The relationship on the plots looks exponential, so we would like to transform the variables and plot logs of variables.

```{r}
qplot(H.Mortality.U5, H.Exp.PerCapita, data = wbi2010, log="xy", xlab="Health Expenditures Per Capita (log)", ylab="Children Under 5 Mortality Rate(log)", main="Health Expenditures vs Mortality Rate")
qplot(H.Mortality.U5, F.GDP.PerCapita, data = wbi2010, log="xy",xlab="GDP Per Capita(log)", ylab="Children Under 5 Mortality Rate(log)", main="GDP vs Mortality Rate")
lm1 <- lm(log(H.Mortality.U5) ~ log(H.Exp.PerCapita),data = wbi2010)
summary(lm1)
```

The relationship now looks linear and confirms that countries with higher Health Expenditures and higher GDP Per Capita have lower mortality rate.

Let's look at trends in Child Mortality Rate between 2000 and 2010 by Region

```{r}
wbi_g<-aggregate(H.Mortality.U5 ~ Region + Date, wbi,sum)
wbi_g$Date2 <- as.character(wbi_g$Date) 
p <- ggplot(wbi_g, aes(Date, H.Mortality.U5, group = Region,  colour = factor(Region))) +
geom_line()+xlab("Year")+ylab("Children Under 5 Mortality Rate")+
ggtitle("Trend in Children Under 5 Mortality Rate")+scale_colour_discrete(name  ="Region")
p
```

The graph shows that the most significant decrease in Child Mortality Rate between 2000 and 2010 happened in Africa.