---
title: "Week 4 Quiz"
output: html_document
---
Q1. Show an appropriate visualization that displays the total number of movies for each decade.

```{r}
library(ggplot2)
data(movies)
genre <- c("Action","Animation", "Comedy", "Drama","Documentary", "Romance", "Short")
ggplot(data = movies, aes(x=year)) + geom_histogram(color="black",fill="white",binwidth=10)+
xlab("Year") + ylab("Movies")+ggtitle("Movies by Decade")
```

Q2. Show the average IMDB user rating for different genres of movies? Has this changed over time?

```{r}
meanrating <- vector(mode = "numeric", length = 0)
for ( i in 1:length(genre)) {
  meanrating[i] <- mean(movies$rating[movies[[genre[i]]]==1])
}

df <- data.frame(movieGenre = genre, movieRating = meanrating )
ggplot(data=df, aes(x=movieGenre, y=movieRating )) + 
    geom_bar(colour="black",fill="white", stat="identity",
             position=position_dodge(),
             size=.3,width = 0.5) +
    xlab("Genre") + ylab("Average Rating") + # Set axis labels
    ggtitle("Average Rating per Genre") 
```

Q3. Is there a relationship between length of movie and movie rating?

The following graph shows that there is no relationship between length of movie and movie rating.

```{r}
m<-movies[movies$length < 500,]
ggplot(data=m, aes(x=length, y=rating)) +
geom_point()
```

Q4. Is there a relationship between length of movie and genre?

The longest movies fall under Action, Romance and Drama category. Note that Short and Animation movies are naturally shorter in length.

```{r}
meanlength <- vector(mode = "numeric", length = 0)
for ( i in 1:length(genre)) {
  meanlength[i] <- mean(movies$length[movies[[genre[i]]]==1])
}
df <- data.frame(movieGenre = genre, movieLength = meanlength )
ggplot(data=df, aes(x=movieGenre, y=movieLength)) + 
    geom_bar(colour="black",fill="white", stat="identity",
             position=position_dodge(),
             size=.3,width = 0.5) +                        # Thinner lines
    xlab("Genre") + ylab("Average Movie Length (in min)") + # Set axis labels
    ggtitle("Average Length per Genre") 
```

Q5. Which other variable best predicts total number of votes that a movie received.

Movies produced in later years receive more votes. This can be attributed to the nature
of internet-based ratings that did not exist until late 90s.

```{r}
ggplot(data=movies, aes(x=year, y=votes)) + geom_point(shape=1) + geom_smooth(method=lm)
```

Number of votes for movies of different lenght shows approximately normal distribution.
Movies of average length receive more votes vs shorter or longer movies.

```{r}
m<-movies[movies$length < 500,]
ggplot(data=m, aes(x=length, y=votes)) + geom_point(shape=1) + geom_smooth(method=lm) 
```

Action movies receive on average more votes than any other genre

```{r}
votes <- vector(mode = "numeric", length = 0)

for ( i in 1:length(genre)) {
  votes[i] <- mean(movies$votes[movies[[genre[i]]]==1])
}
df <- data.frame(movieGenre = genre, movieVotes = votes)
options(scipen=3)
ggplot(data=df, aes(x=movieGenre, y=movieVotes )) + 
    geom_bar(colour="black",fill="white", stat="identity",
             position=position_dodge(),
             size=.3,width = 0.5) +
    xlab("Genre") + ylab("Average Votes") + 
    ggtitle("Average Votes per Genre")
```

 

